% -*- mode: latex; coding: utf-8 -*-
%-----------------------------------------------------------------------------
% pxLST -- User manual and API reference
% Part of P3CTeX. Author: Scvria.
%
% Build from directory tex/ (TEXINPUTS so package is found):
%   set TEXINPUTS=.;./latex;./code;
%   pdflatex -interaction=nonstopmode doc/pxLST.tex
%   pdflatex -interaction=nonstopmode doc/pxLST.tex
%-----------------------------------------------------------------------------
\documentclass[11pt,a4paper]{article}
\PassOptionsToPackage{colorlinks,urlcolor=blue,linkcolor=blue,hyperindex}{hyperref}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{amsmath}
\usepackage{pxLST}

\title{The \texttt{pxLST} package\\
  \large Professional code listing environments for \LaTeX2e}
\author{Scvria\\P3CTeX / UOC PEC Repository}
\date{2026/03/04 -- Version 0.1}

\begin{document}
\maketitle

\begin{abstract}
  \texttt{pxLST} provides a high-level \LaTeX2e\ API for typesetting
  syntax-highlighted source code listings inside styled \texttt{tcolorbox}
  frames.  It wraps the \texttt{listings} and \texttt{tcolorbox} packages
  behind a unified key--value interface (the \texttt{pxLST} key family)
  and ships four visual style renderers based on the Darcula colour palette,
  seven pre-defined language grammars (Java, Prolog, OWL, TAD, YAML,
  Catalan pseudocode, Bash), a named preset system, and a diagnostic probe
  for quality-assurance harnesses.  Full UTF-8 literate support for accented
  Latin characters is included at the global \texttt{lstset} baseline.
\end{abstract}

\tableofcontents

%=============================================================================
\section{Loading the package}
%=============================================================================

\begin{verbatim}
\usepackage{pxLST}
\end{verbatim}

Package options are forwarded to the internal key family \texttt{pxLST}
and are equivalent to calling \verb|\pxLSTsetup| in the preamble.
All options are \texttt{key=value} pairs:

\begin{verbatim}
\usepackage[style=light,language=pxBash]{pxLST}
\end{verbatim}

Defaults are chosen to be sensible for course work and technical reports:
dark Darcula background, Java language, line numbers on the left.  All keys
affect subsequent listings until changed again; the usual \LaTeX\ group
scoping rules apply.

%=============================================================================
\section{Key reference}
%=============================================================================

The \texttt{pxLST} key family contains the following keys.  All of them can be
set at load time via \verb|\usepackage[<keys>]{pxLST}|, globally at any
point via \verb|\pxLSTsetup{<keys>}|, or locally per listing in the
optional \verb|[<options>]| argument of every environment and command.

\begin{center}
\begin{tabular}{llll}
\textbf{Key} & \textbf{Type} & \textbf{Default} & \textbf{Description}
\\\hline
\texttt{style}      & choice  & \texttt{dark}          &
  Visual renderer: \texttt{dark}, \texttt{framed}, \texttt{light},
  \texttt{plain}.\\
\texttt{language}   & tl      & \texttt{pxJava}        &
  Any \verb|\lstdefinelanguage| name; controls grammar and keywords.\\
\texttt{numbers}    & tl      & \texttt{left}          &
  Line number position: \texttt{left}, \texttt{right}, \texttt{none}.\\
\texttt{caption}    & tl      & (empty)                &
  Caption text; stored but not auto-used in \texttt{pxCode}.\\
\texttt{label}      & tl      & (empty)                &
  Cross-reference label; inserted as \verb|\label{lst:<value>}|.\\
\texttt{float}      & tl      & (empty)                &
  Float placement (\texttt{h}, \texttt{t}, \texttt{htbp}\ldots);
  empty~$=$~inline.\\
\texttt{breakable}  & bool    & \texttt{true}          &
  Allow \texttt{tcolorbox} page-break; \texttt{pxCodeBreak} forces true.\\
\texttt{fontsize}   & tl      & \verb|\footnotesize|   &
  Font-size command applied in \texttt{basicstyle}.\\
\texttt{linespread} & tl      & \texttt{1.2}           &
  Numeric factor forwarded to \verb|\linespread| inside \texttt{basicstyle}.\\
\texttt{tabsize}    & int     & \texttt{4}             &
  Tab width in spaces (forwarded to \texttt{listings}).\\
\texttt{preset}     & code    & ---                    &
  Apply a named preset; calls \verb|\pxLSTusepreset{<name>}|.\\
\texttt{probe}      & bool    & \texttt{false}         &
  Emit \texttt{PXLST\_ASSERT:} diagnostics to the log; see~\S\ref{sec:probe}.\\
\texttt{collect}   & bool    & \texttt{true}          &
  If \texttt{true}, append snippets to the code annex; if \texttt{false}, only inline rendering (see~\S\ref{sec:inline-annex}).\\
\texttt{annex-id}   & tl      & (empty)               &
  Scope identifier for the code annex; empty = default document-wide; non-empty = named scope (see~\S\ref{sec:annex-advanced}).\\
\texttt{scope}      & tl      & \texttt{document}     &
  Reserved for future use; per-document scope only in this release.\\
\texttt{title}      & tl      & (empty)               &
  Section title for the code annex when using \verb|\printPxCodeAnnex|.\\
\end{tabular}
\end{center}

\noindent Unknown keys are silently ignored (reserved for future
extension).

\paragraph{Key interaction notes.}
\begin{itemize}
  \item \texttt{style} controls both the listings colour scheme
    (\verb|\lstdefinestyle|) and the tcolorbox frame colours
    (\texttt{colback}, \texttt{colframe}, \texttt{coltext}).
  \item \texttt{language} is independent of \texttt{style}; both
    apply simultaneously to the same listing box.
  \item \texttt{preset} is a \texttt{.code:n} key, so subsequent keys
    override preset values when they appear after \texttt{preset} in
    the option list.  See~\S\ref{sec:presets}.
  \item \texttt{float} and \texttt{breakable} are orthogonal: if
    \texttt{float} is non-empty the listing is wrapped in a
    \texttt{figure} float, and \texttt{breakable} is ignored (floats
    cannot break across pages in standard \LaTeX).
  \item \texttt{caption} and \texttt{label} are stored but are
    \emph{not} automatically used in \texttt{pxCode}; they are used
    by \texttt{pxCodeBox} and \texttt{pxCodeBreak} to build the
    title bar and place \verb|\label|.
\end{itemize}

%=============================================================================
\section{Global configuration --- \texttt{\textbackslash pxLSTsetup}}
%=============================================================================

\begin{verbatim}
\pxLSTsetup{<key>=<value>,...}
\end{verbatim}

Sets package-wide defaults in the \texttt{pxLST} key family.  Usual
\LaTeX\ scoping rules apply, so a setup call inside a group affects only
the listings in that group.  At document level the settings remain in
effect until overridden by another \verb|\pxLSTsetup| call or until a
surrounding group ends.

Example --- select the light style and Bash language for all subsequent
listings in the current scope:

\begin{verbatim}
\pxLSTsetup{style=light, language=pxBash, numbers=none}
\end{verbatim}

%=============================================================================
\section{Environments}
%=============================================================================

All three environments below use the same \texttt{pxlst@configure}
mechanism internally: the optional \verb|[<options>]| argument is
processed first (applying local key overrides), then the display macros
are updated inside a \TeX\ group, so settings revert automatically when
the environment ends.

%-----------------------------------------------------------------------------
\subsection{\texttt{pxCode}}
%-----------------------------------------------------------------------------

\begin{verbatim}
\begin{pxCode}[<options>]{<label>}
  ... code content ...
\end{pxCode}
\end{verbatim}

\begin{itemize}
  \item \texttt{\#1} (optional) --- key--value overrides, local to this
    environment instance.
  \item \texttt{\#2} (mandatory) --- label string; displayed in the title
    bar as \verb|\texttt{\bfseries #2}| and used as the \LaTeX\ label
    anchor \verb|\label{lst:<label>}| when non-empty.
\end{itemize}

The environment wraps a \texttt{listings} \texttt{lstlisting} environment
inside a \texttt{tcolorbox} with \texttt{listing~only} mode, a
\texttt{drop~lifted~shadow} decoration, and a rounded frame (\texttt{arc=.25em}).
It is defined via \verb|\newtcblisting| so that verbatim code inside is
handled correctly.

\paragraph{Example:}

\begin{verbatim}
\begin{pxCode}[language=pxJava, numbers=left]{HelloWorld.java}
public class HelloWorld {
    public static void main(String[] args) {
        System.out.println("Hello, world!");
    }
}
\end{pxCode}
\end{verbatim}

%-----------------------------------------------------------------------------
\subsection{\texttt{pxCodeBox}}
%-----------------------------------------------------------------------------

\begin{verbatim}
\begin{pxCodeBox}[<options>]{<label>}{<title>}{<note>}{<caption>}
  ... code content ...
\end{pxCodeBox}
\end{verbatim}

\begin{itemize}
  \item \texttt{\#1} (optional) --- key--value overrides, local scope.
  \item \texttt{\#2} --- label string; inserted as
    \verb|\label{cdbx:<label>}| in the title bar and used in
    continuation headers on subsequent pages.
  \item \texttt{\#3} --- title text, shown right-aligned in the title bar.
  \item \texttt{\#4} --- short note identifier, shown left-aligned in the
    title bar in a smaller, lighter font.
  \item \texttt{\#5} --- caption text, rendered in the comment panel below
    the listing.
\end{itemize}

\texttt{pxCodeBox} renders a two-region \texttt{tcolorbox}: the main (top)
panel contains the syntax-highlighted listing; the comment (bottom) panel
displays the \texttt{\#5} caption alongside a small codebox anchor
label derived from \texttt{\#2}.  The box is \emph{not} breakable by
default (use \texttt{pxCodeBreak} for long listings).

\paragraph{Example:}

\begin{verbatim}
\begin{pxCodeBox}[style=dark, language=pxJava]
                 {alg:bubbleSort}
                 {BubbleSort.java}
                 {alg}
                 {Classic O($n^2$) comparison sort.}
public void bubbleSort(int[] a) {
    for (int i = 0; i < a.length - 1; i++)
        for (int j = 0; j < a.length - 1 - i; j++)
            if (a[j] > a[j+1]) { int t = a[j]; a[j]=a[j+1]; a[j+1]=t; }
}
\end{pxCodeBox}
\end{verbatim}

%-----------------------------------------------------------------------------
\subsection{\texttt{pxCodeBreak}}
%-----------------------------------------------------------------------------

\begin{verbatim}
\begin{pxCodeBreak}[<options>]{<label>}{<title>}{<note>}{<caption>}
  ... code content ...
\end{pxCodeBreak}
\end{verbatim}

Identical signature to \texttt{pxCodeBox} but forces
\texttt{breakable=true} regardless of any key setting.  Use for long
listings that span multiple pages; the title bar is repeated on
continuation pages with a ``\ldots{}continues from page~$N$'' prefix.

%=============================================================================
\section{File input --- \texttt{\textbackslash pxLSTinput}}
%=============================================================================

\begin{verbatim}
\pxLSTinput[<options>]{<filename>}
\end{verbatim}

Reads an external source file and typesets it as a listing using the
current key settings (with optional local overrides in \texttt{\#1}).
Internally delegates to a \texttt{tcolorbox} that wraps
\verb|\lstinputlisting|, so the full Darcula frame is preserved.

\begin{itemize}
  \item \texttt{\#1} (optional) --- any \texttt{pxLST} key--value pairs.
  \item \texttt{\#2} --- filename, relative to the document's current
    working directory (i.e.\ where \texttt{pdflatex} is invoked from).
\end{itemize}

\paragraph{Example:}

\begin{verbatim}
\pxLSTinput[style=light, language=pxBash]{scripts/build.sh}
\end{verbatim}

Unlike the verbatim environments, \verb|\pxLSTinput| is defined via
\verb|\NewDocumentCommand| and can therefore appear in moving arguments
and macros without special treatment.

%=============================================================================
\section{Inline code and code annex}
\label{sec:inline-annex}
%=============================================================================

This section describes the command that typesets short code snippets
\emph{inline} in the main text and, optionally, collects them for a
consolidated listing at the end of the document.

The \textbf{code annex} is a section at the end of the document where all
code snippets collected with \verb|\pxCodeInline| are listed when you call
\verb|\printPxCodeAnnex|.

\paragraph{Naming: pxCode environment vs.\ inline code command.}
The package provides two different mechanisms involving ``code'':
\begin{itemize}
  \item \textbf{pxCode environment} --- \verb|\begin{pxCode}[<options>]{<label>}| \ldots\ \verb|\end{pxCode}|.
    A single block of verbatim code inside a styled tcolorbox; see~\S\ref{sec:styles}.
  \item \textbf{Inline/annex code command} --- \verb|\pxCodeInline|.
    Takes one or more short code snippets as braced arguments, renders them inline, and can collect them for \verb|\printPxCodeAnnex|.
    Snippets are normal braced material (not verbatim); for long verbatim blocks use the \textbf{pxCode environment}.
\end{itemize}

%-----------------------------------------------------------------------------
\subsection{\texttt{\textbackslash pxCodeInline}}
%-----------------------------------------------------------------------------

\begin{verbatim}
\pxCodeInline[<options>]{<snippet1>}{<snippet2>}...{<snippet8>}
\end{verbatim}

\begin{itemize}
  \item \texttt{\#1} (optional) --- \texttt{pxLST} keys: \texttt{style}, \texttt{language}, \texttt{preset}, \texttt{collect}, \texttt{annex-id}.
    Per-call overrides apply to both the inline rendering and the stored copy for the code annex.
  \item \texttt{\#2}--\texttt{\#9} --- up to \textbf{8} code snippets; each braced argument is one snippet.
    For fewer than 8 snippets, use trailing empty braces, e.g.\ \verb|\pxCodeInline{int x=0;}{return x;}{}{}{}{}{}{}|.
\end{itemize}

Content is \emph{not} verbatim: special characters (\texttt{\#}, \texttt{\%}, \texttt{\&}) must be escaped as in normal \LaTeX.
For full verbatim blocks use \verb|\begin{pxCode}...\end{pxCode}|.

\paragraph{Semantics.}
(1)~\textbf{Inline:} each snippet is rendered as a short inline listing (small tcbox + lstlisting).
(2)~\textbf{Collection:} if \texttt{collect=true}, each snippet (with its effective style/language at call time) is appended to the code annex for the current \texttt{annex-id}.

%-----------------------------------------------------------------------------
\subsection{\texttt{\textbackslash printPxCodeAnnex}}
%-----------------------------------------------------------------------------

\begin{verbatim}
\printPxCodeAnnex[<options>]
\end{verbatim}

Output the collected code snippets at the call site.  Place this command
where the code annex should appear (e.g.\ after \verb|\appendix|).

\begin{itemize}
  \item \texttt{title} --- if non-empty, a \verb|\section*| (or similar) is emitted above the code block.
\end{itemize}

Printing does \emph{not} clear the collection; calling \verb|\printPxCodeAnnex| again outputs the same content.
For multiple named scopes (advanced), see~\S\ref{sec:annex-advanced}.

%-----------------------------------------------------------------------------
\subsection{Keys for inline and annex}
%-----------------------------------------------------------------------------

\begin{center}
\begin{tabular}{llll}
\textbf{Key} & \textbf{Type} & \textbf{Default} & \textbf{Use}\\\hline
\texttt{collect}  & bool & \texttt{true} & Append to code annex; \texttt{false} = inline only.\\
\texttt{annex-id} & tl   & (empty)       & Scope name; empty = default annex (see~\S\ref{sec:annex-advanced}).\\
\texttt{title}    & tl   & (empty)       & Section title for the code annex when using \verb|\printPxCodeAnnex|.\\
\end{tabular}
\end{center}

\verb|\pxLSTsetup| and presets apply: e.g.\ \verb|\pxLSTsetup{style=light}| affects the style used for \verb|\pxCodeInline| and for the annexed code; \verb|\pxCodeInline[preset=minimal, language=pxBash]{echo hello}{}{}{}{}{}{}{}| uses that preset and language for that call and for the stored annex copy.

%-----------------------------------------------------------------------------
\subsubsection{Scoped annex (annex-id) --- advanced}
\label{sec:annex-advanced}
%-----------------------------------------------------------------------------

\paragraph{annex-id.}
If set, only snippets collected with that same \texttt{annex-id} are output when \verb|\printPxCodeAnnex[annex-id=...]| is called; empty = default document-wide scope.

\paragraph{scope.}
Reserved for future use; in this release only per-document scope is supported.

%-----------------------------------------------------------------------------
\subsection{Full example: inline code + code annex}
%-----------------------------------------------------------------------------

The following example uses inline code in the body, then prints the code annex in an appendix.
Interaction with \verb|\pxLSTsetup| and presets is as described above.

\begin{verbatim}
\documentclass{article}
\usepackage{pxLST}

\begin{document}
\section{Exercise}
Implement: \pxCodeInline{int x = 0;}{return x + 1;}{}{}{}{}{}{}.

\pxLSTsetup{style=light}
Elsewhere: \pxCodeInline[language=pxBash]{echo hello}{}{}{}{}{}{}{}.

\appendix
\section*{Annex}
\printPxCodeAnnex[title=Code snippets]
\end{document}
\end{verbatim}

Omitting \verb|\printPxCodeAnnex| does not change the main-text output; annex printing is explicitly requested by the user (opt-in).

%=============================================================================
\section{Style reference}
\label{sec:styles}
%=============================================================================

\texttt{pxLST} provides four visual style renderers, each implemented
as a \verb|\lstdefinestyle| entry paired with corresponding
\texttt{tcolorbox} colour parameters.  All four styles use the Darcula
colour palette defined at package load time.

\begin{center}
\begin{tabular}{llll}
\textbf{Style} & \textbf{Background} & \textbf{Frame} &
  \textbf{Syntax colours}\\\hline
\texttt{dark}   & Darcula dark (\texttt{RGB\{40,44,52\}}) & none &
  Full Darcula palette on dark ground.\\
\texttt{framed} & Darcula dark & left rule (darcula-comment) &
  Full Darcula palette; left rule decorates the gutter.\\
\texttt{light}  & white & none &
  Darcula tones darkened for legibility on white.\\
\texttt{plain}  & white & none &
  Monochrome: keywords bold black, comments italic grey.\\
\end{tabular}
\end{center}

\paragraph{tcolorbox parameters per style.}

\begin{center}
\begin{tabular}{lllll}
\textbf{Style} & \textbf{colback} & \textbf{colframe} &
  \textbf{coltext} & \textbf{arc}\\\hline
\texttt{dark}   & \texttt{darcula-bg!90!} & \texttt{darcula-bg!60!}     & \texttt{darcula-fg} & \texttt{.25em}\\
\texttt{framed} & \texttt{darcula-bg!90!} & \texttt{darcula-comment!60!}& \texttt{darcula-fg} & \texttt{.25em}\\
\texttt{light}  & \texttt{white}          & \texttt{gray!25!}           & \texttt{black}      & \texttt{.25em}\\
\texttt{plain}  & \texttt{white}          & \texttt{gray!15!}           & \texttt{black}      & \texttt{0pt}\\
\end{tabular}
\end{center}

\paragraph{Switching styles.}

\begin{verbatim}
% Globally:
\pxLSTsetup{style=framed}

% Locally in one listing:
\begin{pxCode}[style=light]{MyFile.java}
  ...
\end{pxCode}
\end{verbatim}

The \texttt{dark} and \texttt{framed} styles are recommended for
standalone course submissions and printable reports where a clear
visual boundary is desired.  The \texttt{light} style suits documents
that mix code with running prose and share the page background with
text blocks.  The \texttt{plain} style is appropriate for black-and-white
printing where colour ink should be avoided.

%=============================================================================
\section{Language reference}
\label{sec:languages}
%=============================================================================

Seven language grammars are defined in \texttt{pxLST.code.tex} via
\verb|\lstdefinelanguage|.  They are available regardless of the active
style.

\begin{center}
\begin{tabular}{lll}
\textbf{Language key} & \textbf{Based on} & \textbf{Typical use}\\\hline
\texttt{pxJava}    & Java  & Java source, annotations (\texttt{@}), generics.\\
\texttt{pxProlog}  & Prolog & Logic programming, facts, rules, DCG clauses.\\
\texttt{pxOWL}     & --- & Manchester OWL ontology syntax.\\
\texttt{pxTAD}     & --- & Abstract data type (TAD) specifications.\\
\texttt{pxYAML}    & --- & YAML configuration files.\\
\texttt{algoritme} & --- & Catalan pseudocode with rendered math symbols.\\
\texttt{pxBash}    & bash & Shell scripts, command-line listings.\\
\end{tabular}
\end{center}

\paragraph{UTF-8 literate map.}
The global \verb|\lstset| baseline loaded by \texttt{pxLST} includes a
comprehensive UTF-8 literate mapping for accented characters: acute,
grave, umlaut, and circumflex accents are mapped for all letters
\texttt{a--z} and \texttt{A--Z}, plus the digraphs
\texttt{\c{c}}, \texttt{\c{C}}, \texttt{\~n}, \texttt{\~N},
\texttt{\textquestiondown}, and \texttt{\textexclamdown}.
This means Catalan, Spanish, and French source files containing accented
identifiers or comments are typeset correctly without any manual literate
entry.

\paragraph{The \texttt{algoritme} language.}
This language is designed for Catalan pseudocode used in UOC coursework.
It maps common ASCII sequences to proper mathematical symbols:

\begin{center}
\begin{tabular}{ll}
\textbf{Input} & \textbf{Output}\\\hline
\texttt{<{-{}-}}\quad (three chars) & $\leftarrow$ (assignment arrow)\\
\texttt{<=}    & $\leq$\\
\texttt{>=}    & $\geq$\\
\texttt{!=}    & $\neq$\\
\texttt{*}     & $\times$\\
\end{tabular}
\end{center}

Keywords (\texttt{Funcio}, \texttt{Fi}, \texttt{Si}, \texttt{Llavors},
\texttt{Sinó}, \texttt{Mentre}, \texttt{Per}, \texttt{Retornar}\ldots)
are rendered bold in the Darcula keyword colour; case-sensitive matching
is active.

%=============================================================================
\section{Preset system}
\label{sec:presets}
%=============================================================================

A \emph{preset} is a named bundle of key--value settings that can be
saved once and applied to any number of listings.  Presets survive group
boundaries when saved, but their \emph{effect} is local to the scope in
which they are applied.

\subsection{Preset commands}

\begin{itemize}
  \item \verb|\pxLSTsavepreset{<name>}{<option-string>}|\\
    Stores the option string under the given name.  If a preset with the
    same name already exists it is silently overwritten.

  \item \verb|\pxLSTusepreset{<name>}|\\
    Applies the preset in the current scope, equivalent to calling
    \verb|\pxLSTsetup| with the stored option string.  If the preset is
    not found, a warning is written to the log and the call is ignored.

  \item \texttt{preset=<name>} key\\
    The \texttt{preset} key can be used inside any option list.  Because
    keys are processed in source order, keys appearing \emph{after}
    \texttt{preset} override preset values:

\begin{verbatim}
% 'darcula' sets language=pxJava; the subsequent key overrides it.
\pxLSTsetup{preset=darcula, language=pxBash}
\end{verbatim}

\end{itemize}

\subsection{Built-in presets}

Three generic presets are always available (registered at package load
time):

\begin{center}
\begin{tabular}{lp{.55\linewidth}}
\textbf{Preset} & \textbf{Key bundle}\\\hline
\texttt{darcula}  &
  \texttt{style=dark, language=pxJava, numbers=left, breakable=true}\\
\texttt{minimal}  &
  \texttt{style=plain, language=pxJava, numbers=none}\\
\texttt{academic} &
  \texttt{style=light, language=pxJava, numbers=left, fontsize=\textbackslash small}\\
\end{tabular}
\end{center}

\subsection{P3CTeX branded presets}

When \texttt{pxLST} is loaded via the \texttt{P3CTeX} document class,
three additional branded presets are registered automatically before
option processing, so they can be used in \verb|\usepackage| options:

\begin{verbatim}
\usepackage[preset=p3c-code]{pxLST}
\end{verbatim}

\begin{center}
\begin{tabular}{lp{.58\linewidth}}
\textbf{Preset} & \textbf{Key bundle}\\\hline
\texttt{p3c-code}  &
  \texttt{style=dark, language=pxJava, numbers=left, breakable=true}\\
\texttt{p3c-alg}   &
  \texttt{style=framed, language=algoritme, numbers=left, breakable=false}\\
\texttt{p3c-plain} &
  \texttt{style=plain, language=pxJava, numbers=none, breakable=false}\\
\end{tabular}
\end{center}

\subsection{Usage example}

\begin{verbatim}
% Define a custom preset:
\pxLSTsavepreset{myprolog}{style=framed, language=pxProlog, numbers=none}

% Apply it globally:
\pxLSTusepreset{myprolog}

% Or locally:
\begin{pxCode}[preset=myprolog]{family.pl}
  parent(tom, bob). parent(tom, liz).
  ancestor(X,Y) :- parent(X,Y).
\end{pxCode}

% Override one key after preset:
\begin{pxCode}[preset=myprolog, numbers=left]{rules.pl}
  ...
\end{pxCode}
\end{verbatim}

%=============================================================================
\section{pxCORE integration}
\label{sec:pxcore}
%=============================================================================

\texttt{pxLST} is an opt-in module of \texttt{pxCORE}, the P3CTeX
core package.  By default, \texttt{pxCORE} does \emph{not} load
\texttt{pxLST} (initial value of the \texttt{LST} module key is
\texttt{false}).

\paragraph{Explicit opt-in.}
\begin{verbatim}
\usepackage[LST]{pxCORE}
\end{verbatim}
This sets \texttt{LST=true} and causes \verb|\RequirePackage{pxLST}| to
be executed during \texttt{pxCORE} loading.

\paragraph{Bundle (all modules).}
\begin{verbatim}
\usepackage[default]{pxCORE}
\end{verbatim}
The \texttt{default} meta-key activates all modules:
\texttt{PRP}, \texttt{UML}, \texttt{SRC}, \texttt{TAB}, and \texttt{LST}.
This is the recommended setup when using the full P3CTeX workflow.

\paragraph{Module key summary.}

\begin{center}
\begin{tabular}{lll}
\textbf{Usage} & \textbf{LST loaded?} & \textbf{Note}\\\hline
\verb|\usepackage{pxCORE}|            & No  & Default (opt-in required).\\
\verb|\usepackage[LST]{pxCORE}|       & Yes & Explicit opt-in.\\
\verb|\usepackage[default]{pxCORE}|   & Yes & All modules bundle.\\
\end{tabular}
\end{center}

%=============================================================================
\section{Diagnostic probe}
\label{sec:probe}
%=============================================================================

\begin{verbatim}
\pxLSTsetup{probe=true}
\end{verbatim}

When \texttt{probe} is set to \texttt{true}, every call to
\verb|\pxLSTsetup| (and every environment invocation) writes a block
of diagnostic lines to the \LaTeX\ log via \verb|\iow_term:x|.  These
lines start with the prefix \texttt{PXLST\_ASSERT:} and are intended
for automated QA harnesses that parse the log file.

\paragraph{Asserted values.}

\begin{center}
\begin{tabular}{ll}
\textbf{Log key} & \textbf{Source variable}\\\hline
\texttt{PXLST\_ASSERT: STYLE}    & \verb|\l_pxlst_style_tl| (e.g.\ \texttt{dark})\\
\texttt{PXLST\_ASSERT: LANGUAGE} & \verb|\l_pxlst_language_tl| (e.g.\ \texttt{pxJava})\\
\texttt{PXLST\_ASSERT: NUMBERS}  & \verb|\l_pxlst_numbers_tl| (e.g.\ \texttt{left})\\
\texttt{PXLST\_ASSERT: COLBACK}  & \verb|\pxlst@colback@v| (e.g.\ \texttt{darcula-bg!90!})\\
\texttt{PXLST\_ASSERT: LSTYLE}   & \verb|\pxlst@lstyle@v| (e.g.\ \texttt{pxlst@dark})\\
\end{tabular}
\end{center}

Probe output has no effect on the typeset document.  Disable it again
with \verb|\pxLSTsetup{probe=false}| once debugging is complete.

%=============================================================================
\section{Known limitations (v0.1)}
\label{sec:limits}
%=============================================================================

The following items are explicitly outside the scope of the current
release and are deferred to a future sprint.

\begin{itemize}
  \item \textbf{No \texttt{minted} backend.}  \texttt{pxLST} v0.1
    uses the \texttt{listings} package exclusively.  A Pygments-based
    (\texttt{minted}) rendering backend is a v0.2 candidate.

  \item \textbf{No dedicated listing float counter.}  The \texttt{float}
    key wraps listings in a \texttt{figure} environment.  A separate
    \texttt{listing} float counter (compatible with \verb|\listoflistings|)
    is deferred.

  \item \textbf{No automatic language detection.}  \verb|\pxLSTinput|
    does not infer the language from the file extension; the
    \texttt{language} key must be set explicitly.

  \item \textbf{No side-by-side listing environments.}  Two-column
    code comparisons (source vs.\ output, before vs.\ after) are not
    provided in this version.

  \item \textbf{No per-environment keyword extension.}  There is no
    high-level key to add \texttt{morekeywords} at use time beyond the
    fixed language definition; users who need additional keywords must
    use raw \verb|\lstset| calls.

  \item \textbf{No \texttt{pxANX} integration.}  Document-local code
    annex is supported via \verb|\printPxCodeAnnex| (see~\S\ref{sec:inline-annex}).  Automatic
    population of a list-of-listings via the P3CTeX annex subsystem
    (pxANX) is deferred to a future sprint.
\end{itemize}

%=============================================================================
\section{Implementation notes}
%=============================================================================

This section gives a brief account of the internal architecture for
maintainers and advanced users.

\subsection*{File organisation}

\texttt{pxLST} is split across two files:
\begin{itemize}
  \item \texttt{tex/latex/pxLST.sty} --- public \LaTeX2e\ API:
    \verb|\RequirePackage| block, key-processing bootstrap,
    public commands (\verb|\pxLSTsetup|, \verb|\pxLSTsavepreset|,
    \verb|\pxLSTusepreset|, \verb|\pxLSTinput|, \verb|\pxCodeInline|,
    \verb|\printPxCodeAnnex|), and the three
    \verb|\newtcblisting| environment declarations.
  \item \texttt{tex/code/pxLST.code.tex} --- expl3 internals: variable
    declarations, key definitions, Darcula colour palette, display macro
    initialisations, global \verb|\lstset| baseline, four style
    definitions, seven language definitions, preset mechanism, probe,
    and \verb|\pxLST_input_file:nn|.
\end{itemize}

The \texttt{.sty} inputs the \texttt{.code.tex} file via
\verb|\file_input:n{pxLST.code.tex}|, so both files must be on the
\texttt{TEXINPUTS} path.

\subsection*{Dynamic option mechanism}

Since \verb|\newtcblisting| environments must handle verbatim content,
they cannot call \verb|\pxLSTsetup| directly inside their option
argument.  Instead, a PGF key \texttt{/tcb/pxlst@configure} is defined
in \texttt{.code.tex}:

\begin{verbatim}
\pgfkeys{/tcb/pxlst@configure/.code n args={1}{
  \pxLST_set_options:n { #1 }
  \__pxlst_set_style_vars:
  \tcbset{
    colback  = \pxlst@colback@v,
    colframe = \pxlst@colframe@v,
    ...
  }
}}
\end{verbatim}

Each \verb|\newtcblisting| passes \verb+pxlst@configure={#1}+
as its first tcolorbox option, which triggers the key processing chain
before the box is opened.  The display macros (\verb|\pxlst@*@v|) are
set via \verb|\def| and therefore restore automatically when the
group closed by \verb|\end{pxCode}| (etc.) unwinds.

\subsection*{Dependency stack}

\texttt{pxLST.sty} loads the following packages in order:
\texttt{expl3}, \texttt{xparse}, \texttt{l3keys2e},
\texttt{listings}, \texttt{listingsutf8}, \texttt{xcolor} (with the
\texttt{table} option), \texttt{tcolorbox}.
The tcolorbox libraries \texttt{listings}, \texttt{listingsutf8},
\texttt{breakable}, \texttt{xparse}, and \texttt{skins} are loaded via
\verb|\tcbuselibrary|.  The \texttt{float} package and \texttt{amssymb}
are \emph{not} required: float wrapping uses the standard \texttt{figure}
environment, and math symbols in the \texttt{algoritme} language are
rendered inline via standard \LaTeX\ math mode.

\end{document}
