% -*- mode: latex; coding: utf-8 -*-
%-----------------------------------------------------------------------------
% pxTAB -- User manual and API reference
% Part of P3CTeX. Author: Scvria.
%
% Build from directory tex/ (TEXINPUTS so package is found):
%   set TEXINPUTS=.;./latex;./code;
%   pdflatex -interaction=nonstopmode doc/pxTAB.tex
%   makeindex -s gind.ist -o pxTAB.ind pxTAB.idx
%   pdflatex -interaction=nonstopmode doc/pxTAB.tex
%-----------------------------------------------------------------------------
\documentclass[11pt,a4paper]{article}
\PassOptionsToPackage{colorlinks,urlcolor=blue,linkcolor=blue,hyperindex}{hyperref}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{amsmath}
\usepackage{pxTAB}

\title{The \texttt{pxTAB} package\\
  \large Predefined table styles and list-driven tables for \LaTeX2e}
\author{Scvria\\P3CTeX / UOC PEC Repository}
\date{2026/03/02 -- Version 0.1}

\begin{document}
\maketitle

\begin{abstract}
  \texttt{pxTAB} provides a small \LaTeX2e\ API to build tables from
  comma-separated rows or list-style data without writing raw
  \texttt{tabular} environments.  It offers a handful of predefined
  visual styles (plain, header, striped, boxed, grid) and a set of
  composed header styles (header-boxed, header-grid, header-striped),
  together with key--value options for consistent layout control.
\end{abstract}

\tableofcontents

\section{Loading the package}

\begin{verbatim}
\usepackage{pxTAB}
\end{verbatim}

Package options are forwarded to the internal key family \texttt{pxTAB}
and are equivalent to calling \verb|\pxTABsetup| in the preamble:

\begin{verbatim}
\usepackage[style=header,headerbgcolor=blue!15]{pxTAB}
\end{verbatim}

\section{Configuration (keys)}

The \texttt{pxTAB} key family controls both visual style and parsing
behaviour.  Keys can be set at load time or via:

\begin{verbatim}
\pxTABsetup{<key>=<value>,...}
\end{verbatim}

Defaults are chosen to be reasonable for course work and reports.
All keys affect subsequent tables until changed again (usual \LaTeX\
scoping rules apply).

\subsection*{Style and colours}

\begin{itemize}
  \item \texttt{style} (choice, default \texttt{plain})\\
    Selects the high-level style:
    \texttt{plain}, \texttt{header}, \texttt{striped},
    \texttt{boxed}, \texttt{grid}, or one of the composed
    header styles \texttt{header-boxed}, \texttt{header-grid},
    \texttt{header-striped}.  The composed styles behave like
    the corresponding base style (boxed, grid, striped) with
    an additional coloured header row.

  \item \texttt{headerbgcolor} (token list, default \texttt{blue!15})\\
    Background colour of the header row in \texttt{header} style.

  \item \texttt{headertextcolor} (token list, default \texttt{blue!70!black})\\
    Text colour used in the header row.

  \item \texttt{altrowbgcolor} (token list, default \texttt{blue!5})\\
    Background colour for alternating rows in \texttt{striped} style.

  \item \texttt{altrowcolor} (meta)\\
    Alias for \texttt{altrowbgcolor}.

  \item \texttt{altrowtextcolor} (token list, default empty)\\
    Text colour used on alternating rows in \texttt{striped} style.

  \item \texttt{headerfont} (token list, default \verb|\bfseries|)\\
    Font-switch applied to the header row (e.g.\ \verb|\sffamily\bfseries|).

  \item \texttt{bodyfont} (token list, default empty)\\
    Font-switch applied to body rows.

  \item \texttt{rulecolor} (token list, default \texttt{black!55})\\
    Colour used for horizontal and vertical rules
    (applied via \verb|\arrayrulecolor|).

  \item \texttt{rulewidth} (dimension, default \texttt{0.4pt})\\
    Thickness of rules (applied via \verb|\arrayrulewidth|).
\end{itemize}

\subsection*{Style shortcuts}

The following meta keys select a style without the \texttt{style=}
prefix.  They take no argument and can be mixed freely with other
keys:

\begin{verbatim}
\pxTABsetup{header-striped, headerbgcolor=teal!20}
\end{verbatim}

\begin{itemize}
  \item \texttt{striped} \quad $\equiv$ \texttt{style=striped}
  \item \texttt{boxed} \quad $\equiv$ \texttt{style=boxed}
  \item \texttt{grid} \quad $\equiv$ \texttt{style=grid}
  \item \texttt{header} \quad $\equiv$ \texttt{style=header}
  \item \texttt{header-striped} \quad $\equiv$ \texttt{style=header-striped}
  \item \texttt{header-boxed} \quad $\equiv$ \texttt{style=header-boxed}
  \item \texttt{header-grid} \quad $\equiv$ \texttt{style=header-grid}
\end{itemize}

\subsection*{Quick header setup}

These convenience keys combine a style selection with a header
background colour in a single key.  The colour argument follows
standard \texttt{xcolor} syntax.  Note that \texttt{headertextcolor}
and \texttt{headerfont} retain their current values (they are not
reset by these keys).

\begin{itemize}
  \item \texttt{setheader=<colour>}\\
    $\equiv$ \texttt{style=header, headerbgcolor=<colour>}
  \item \texttt{setheader-striped=<colour>}\\
    $\equiv$ \texttt{style=header-striped, headerbgcolor=<colour>}
  \item \texttt{setheader-boxed=<colour>}\\
    $\equiv$ \texttt{style=header-boxed, headerbgcolor=<colour>}
  \item \texttt{setheader-grid=<colour>}\\
    $\equiv$ \texttt{style=header-grid, headerbgcolor=<colour>}
\end{itemize}

\noindent Example:

\begin{verbatim}
\pxTable[setheader=teal!20]{Name, Grade}{Alice, 9}
\end{verbatim}

\subsection*{Geometry and layout}

\begin{itemize}
  \item \texttt{tabcolsep} (dimension, default \texttt{6pt})\\
    Column separation (mirrors the standard \verb|\tabcolsep| parameter).

  \item \texttt{arraystretch} (token list, default \texttt{1.2})\\
    Row stretch applied via \verb|\arraystretch|.

  \item \texttt{rowstretch} (meta)\\
    Alias for \texttt{arraystretch}.

  \item \texttt{width} (code, default \texttt{natural})\\
    Width of the table.  The default \texttt{natural} keeps the table
    at its natural (fit-content) width; tables are \emph{not} stretched
    to page width unless you explicitly set \texttt{width} or
    \texttt{tabwidth}.  The special values \texttt{linewidth} and
    \texttt{line-width} are treated as \verb|\linewidth|; any other
    value is used as-is (e.g.\ \verb|.8\linewidth| or \verb|120pt|).

  \item \texttt{tabwidth} (meta)\\
    Alias for \texttt{width}.

  \item \texttt{centering} (boolean, default \texttt{false})\\
    When \texttt{true}, inserts a \verb|\centering| before the internal
    \texttt{tabular} or \texttt{tabular*}.  This only changes the
    horizontal alignment of the table block (in a paragraph or inside a
    \texttt{table} float); it does not modify the computed width.  In
    particular, \verb|width=\linewidth| still produces a full-width
    table, but the block is explicitly centred.  The key can be set
    globally with \verb|\pxTABsetup| or per table in the optional
    \verb|[<options>]| argument of the commands below.

  \item \texttt{fontsize} (token list, default empty)\\
    Optional font-size command (such as \verb|\small| or \verb|\footnotesize|)
    applied to the table.

  \item \texttt{size} (meta)\\
    Alias for \texttt{fontsize}.

  \item \texttt{align} (choice, default \texttt{c})\\
    When \texttt{colspec} is not provided, controls automatic column
    alignment: \texttt{l}, \texttt{c}, or \texttt{r}.

  \item \texttt{alignment} (meta)\\
    Alias for \texttt{align}.

  \item \texttt{colspec} (token list, default empty)\\
    Explicit \texttt{tabular} column specification (e.g.\ \texttt{ll},
    \texttt{|c|c|}).  When omitted, \texttt{pxTAB} builds one
    automatically from \texttt{align}.  Simple \texttt{l}/\texttt{c}/\texttt{r}
    specifications are checked against the inferred column count; on
    mismatch a warning is issued and an automatic specification is used.

\end{itemize}

\subsection*{Floats and captions}

The following keys enable a lightweight float wrapper around any
\texttt{pxTAB} table command.  They can be set at package level or
passed in the per-table options:

\begin{itemize}
  \item \texttt{caption} (token list, default empty)\\
    When non-empty, the table body is wrapped in a \texttt{table}
    environment and \verb|\caption{<caption>}| is emitted above the
    tabular.

  \item \texttt{label} (token list, default empty)\\
    Optional label placed immediately after the caption (or, if there is
    no caption, at the top of the float body).  A recommended convention
    is to use prefixes such as \texttt{tab:grades} so that
    \verb|\ref{tab:grades}| yields the table number in the usual way.

  \item \texttt{float} (token list, default \texttt{H})\\
    Placement specifier forwarded to the \texttt{table} environment
    (e.g.\ \texttt{htbp}).  The default \texttt{H} (from the
    \texttt{float} package, loaded automatically by \texttt{pxTAB})
    pins the table at the exact source position when the float wrapper
    is activated by \texttt{caption}, \texttt{label}, or a non-empty
    \texttt{float} value.
\end{itemize}

Any non-empty combination of \texttt{caption}, \texttt{label}, or
\texttt{float} activates the float wrapper; otherwise tables are
generated in-place without a surrounding \texttt{table} environment.
When both \texttt{centering=true} and float keys are used, the
\verb|\centering| is inserted inside the \texttt{table} environment,
before the caption and the tabular body.

\subsection*{Parsing}

\begin{itemize}
  \item \texttt{rowsep} (token list, default \texttt{;})\\
    Separator between rows when using list-style data.

  \item \texttt{cellsep} (token list, default \texttt{,})\\
    Separator between cells inside each row.

  \item \texttt{probe} (boolean, default \texttt{false})\\
    When \texttt{true}, writes diagnostic information to the \LaTeX\ log
    file during table rendering (style name, colspec, visual settings,
    etc.).  Intended for debugging and development; no effect on
    output.  QA harnesses may override the internal probe hook for
    automated assertions.

  \item \texttt{unknown} (code)\\
    Unknown keys are currently ignored silently (reserved for future
    extension).
\end{itemize}

\subsection*{Paragraph cells}

By default \texttt{pxTAB} produces \emph{inline} columns
(\texttt{l}/\texttt{c}/\texttt{r}), suitable for short labels and
numbers.  Paragraph mode enables fixed-width columns that wrap
multi-line content, mirroring the equal-width paragraph layout
previously available through \texttt{pxTBL}.

\begin{itemize}
  \item \texttt{cellmode} (choice, default \texttt{inline})\\
    Controls the column type used by the automatic colspec builder:
    \begin{itemize}
      \item \texttt{inline} --- standard behaviour;
            columns are simple alignment columns derived from
            \texttt{align} (\texttt{l}/\texttt{c}/\texttt{r}).
      \item \texttt{paragraph} --- each column becomes a paragraph
            column (\texttt{p\{...\}} or \texttt{m\{...\}}) with
            equal width, so cell content wraps across multiple lines.
    \end{itemize}
    Paragraph mode is \emph{opt-in}; existing documents that do not
    set this key are unaffected.

  \item \texttt{paragraphcells} (boolean, default \texttt{true} when used)\\
    Convenience alias:
    \texttt{paragraphcells=true} is equivalent to
    \texttt{cellmode=paragraph}, and
    \texttt{paragraphcells=false} to \texttt{cellmode=inline}.
    When used without an explicit value (bare
    \verb|\pxTABsetup{paragraphcells}|) it defaults to \texttt{true}.

  \item \texttt{paragraphcoltype} (choice, default \texttt{m})\\
    Selects the underlying \LaTeX\ column type used in paragraph mode:
    \begin{itemize}
      \item \texttt{p} --- \texttt{p\{<width>\}} columns; cell content
            is top-aligned within the cell box.
      \item \texttt{m} --- \texttt{m\{<width>\}} columns (requires the
            \texttt{array} package, loaded automatically by
            \texttt{pxTAB}); cell content is vertically centred
            within the cell box.
    \end{itemize}
    This key is ignored when \texttt{cellmode=inline}.

  \item \texttt{paragraph} (meta, no argument)\\
    Convenience alias for \texttt{cellmode=paragraph}.
    Example: \verb|\pxTABsetup{paragraph}|.

  \item \texttt{verbose} (meta, no argument)\\
    Legacy alias for \texttt{cellmode=paragraph}, retained for
    backward compatibility.  Prefer \texttt{paragraph} or the
    explicit \texttt{cellmode=paragraph} form.
\end{itemize}

\paragraph{Effective width.}
When paragraph mode is active the per-column width is computed as
\[
  w_{\text{col}} = \frac{W_{\text{eff}}}{n_{\text{cols}}}
\]
where $W_{\text{eff}}$ is the \emph{effective table width}:
\begin{itemize}
  \item If \texttt{width} (or \texttt{tabwidth}) is set to an explicit
    dimension (e.g.\ \verb|0.8\linewidth|, \texttt{120pt}), that value
    is used directly as $W_{\text{eff}}$.
  \item If \texttt{width=natural} (the default), paragraph mode uses
    \verb|\linewidth| as $W_{\text{eff}}$, since paragraph columns
    require an explicit dimension.  The resulting table occupies the
    full text width by default.
\end{itemize}
Horizontal alignment inside each paragraph cell is governed by the
\texttt{align} key: \texttt{l}~maps to \verb|\raggedright|,
\texttt{c}~to \verb|\centering|, and \texttt{r}~to \verb|\raggedleft|.

\paragraph{Interaction with \texttt{colspec}.}
When the user provides an explicit \texttt{colspec}, it is used as-is
regardless of \texttt{cellmode}.  In other words, \texttt{cellmode=paragraph}
only affects the \emph{automatic} colspec builder; advanced users who
need heterogeneous column widths or mixed column types should set
\texttt{colspec} directly.

\section{Presets}

A \emph{preset} is a named bundle of key--value settings that can be
saved once and applied to any number of tables.  Presets are stored
globally (they survive group boundaries), but their \emph{effect} is
local to the current scope, following normal \LaTeX\ scoping rules.

\subsection{Preset commands}

\begin{itemize}
  \item \verb|\pxTABsavepreset{<name>}{<keys>}|\\
    Stores the key--value string \verb|<keys>| under the name
    \verb|<name>|.  If a preset with the same name already exists it is
    silently overwritten.

  \item \verb|\pxTABusepreset{<name>}|\\
    Applies the preset \verb|<name>| in the current scope (equivalent
    to calling \verb|\pxTABsetup| with the stored keys).  If the preset
    is not defined, a warning is written to the log and the call is
    ignored.

  \item \texttt{preset} key\\
    The \texttt{preset} key can be used inside any per-table option
    list.  Because \texttt{l3keys} processes keys in source order, keys
    appearing \emph{after} \texttt{preset} override preset values:

\begin{verbatim}
\pxTable[preset=formal, headerbgcolor=red!20]
  {Name, Grade}{Alice, 9}
\end{verbatim}

    Here \texttt{formal} sets \texttt{headerbgcolor=blue!15}, but the
    explicit \texttt{headerbgcolor=red!20} overrides it.
\end{itemize}

\subsection{Built-in presets}

Three generic presets are always available (registered at package load
time):

\begin{center}
\begin{tabular}{llp{.48\linewidth}}
\textbf{Preset} & \textbf{Style} & \textbf{Key bundle}\\\hline
\texttt{formal} & \texttt{header} &
  \texttt{headerbgcolor=blue!15},
  \texttt{headertextcolor=blue!70!black},
  \texttt{headerfont=\textbackslash sffamily\textbackslash bfseries},
  \texttt{centering=true}, \texttt{arraystretch=1.3}\\
\texttt{zebra} & \texttt{header-striped} &
  \texttt{headerbgcolor=blue!15},
  \texttt{headertextcolor=blue!70!black},
  \texttt{altrowbgcolor=blue!5},
  \texttt{headerfont=\textbackslash bfseries},
  \texttt{centering=true}\\
\texttt{minimal} & \texttt{plain} &
  \texttt{arraystretch=1.0}, \texttt{tabcolsep=4pt}\\
\end{tabular}
\end{center}

\subsection{Usage example}

\begin{verbatim}
% Define a custom preset:
\pxTABsavepreset{my-rubric}{
  style=header-grid,
  headerbgcolor=teal!30,
  cellmode=paragraph,
  width=\linewidth,
  centering=true
}

% Use it:
\pxTable[preset=my-rubric]{Criterion, Comment}
  {Clarity, Clear and well-structured}
  {Depth, Shows critical analysis}

% Override one key:
\pxTable[preset=my-rubric, headerbgcolor=orange!20]
  {Criterion, Comment}
  {Clarity, Clear and well-structured}

% Apply a built-in preset via \pxTABusepreset:
\begingroup
  \pxTABusepreset{formal}
  \pxTable{Name, Grade}{Alice, 9}{Bob, 7}
\endgroup
\end{verbatim}

\section{Styles by example}

This section illustrates the built-in styles using a small example
dataset.  In all cases we use:

\begin{verbatim}
\pxTABsetup{tabwidth=.8\linewidth}
\end{verbatim}

\subsection{Plain}

\begin{verbatim}
\pxTABsetup{style=plain,tabwidth=.8\linewidth}
\pxTable{Name, Age, City}
        {Alice, 20, Barcelona}
        {Bob, 22, Madrid}
\end{verbatim}

\pxTABsetup{style=plain,tabwidth=.8\linewidth}
\pxTable{Name, Age, City}
        {Alice, 20, Barcelona}
        {Bob, 22, Madrid}

\subsection{Header}

\begin{verbatim}
\pxTABsetup
  {
    style=header,
    headerbgcolor=blue!15,
    headertextcolor=blue!70!black,
    headerfont=\sffamily\bfseries
  }
\pxTable{Name, Age, City}
        {Alice, 20, Barcelona}
        {Bob, 22, Madrid}
\end{verbatim}

\pxTABsetup
  {
    style=header,
    headerbgcolor=blue!15,
    headertextcolor=blue!70!black,
    headerfont=\sffamily\bfseries
  }
\pxTable{Name, Age, City}
        {Alice, 20, Barcelona}
        {Bob, 22, Madrid}

\subsection{Striped}

\begin{verbatim}
\pxTABsetup
  {
    style=striped,
    altrowbgcolor=gray!10,
    altrowtextcolor=black
  }
\pxTable{Item, Quantity}
        {Apples, 3}
        {Oranges, 5}
        {Bananas, 2}

\end{verbatim}

\pxTABsetup
  {
    style=striped,
    altrowbgcolor=gray!10,
    altrowtextcolor=black
  }
\pxTable{Item, Quantity}
        {Apples, 3}
        {Oranges, 5}
        {Bananas, 2}

\noindent
In the \texttt{striped}, \texttt{header-striped}, and
\texttt{header-grid} styles, row backgrounds are applied with
\verb|\rowcolor| so that the colour fills the \emph{entire cell area}
across all columns---including any wrapped lines in paragraph mode.
This holds for both natural-width and explicit-width tables.

\subsection{Boxed}

\begin{verbatim}
\pxTABsetup{style=boxed}
\pxTableFromList
  {Code, Description;
   PX1, First code;
   PX2, Second code}
\end{verbatim}

\pxTABsetup{style=boxed}
\pxTableFromList
  {Code, Description;
   PX1, First code;
   PX2, Second code}

\subsection{Grid}

\begin{verbatim}
\pxTABsetup{style=grid}
\pxTableMatrix{2}{3}
  {m11,m12,m13,
   m21,m22,m23}
\end{verbatim}

\pxTABsetup{style=grid}
\pxTableMatrix{2}{3}
  {m11,m12,m13,
   m21,m22,m23}

\subsection{Composed header styles}

The composed styles combine the header row styling with the boxed,
grid, or striped variants so that common table layouts can be selected
with a single option.  The following examples use the same small
dataset.

\subsubsection*{\texttt{style=header-boxed}}

\begin{verbatim}
\pxTABsetup
  {
    style=header-boxed,
    headerbgcolor=blue!15,
    headertextcolor=blue!70!black,
    headerfont=\sffamily\bfseries,
    tabwidth=.8\linewidth
  }
\pxTable{Name, Age, City}
        {Alice, 20, Barcelona}
        {Bob, 22, Madrid}
\end{verbatim}

\pxTABsetup
  {
    style=header-boxed,
    headerbgcolor=blue!15,
    headertextcolor=blue!70!black,
    headerfont=\sffamily\bfseries,
    tabwidth=.8\linewidth
  }
\pxTable{Name, Age, City}
        {Alice, 20, Barcelona}
        {Bob, 22, Madrid}

\subsubsection*{\texttt{style=header-grid}}

\begin{verbatim}
\pxTABsetup
  {
    style=header-grid,
    headerbgcolor=blue!15,
    headertextcolor=blue!70!black,
    headerfont=\sffamily\bfseries,
    tabwidth=.8\linewidth
  }
\pxTable{Name, Age, City}
        {Alice, 20, Barcelona}
        {Bob, 22, Madrid}
\end{verbatim}

\pxTABsetup
  {
    style=header-grid,
    headerbgcolor=blue!15,
    headertextcolor=blue!70!black,
    headerfont=\sffamily\bfseries,
    tabwidth=.8\linewidth
  }
\pxTable{Name, Age, City}
        {Alice, 20, Barcelona}
        {Bob, 22, Madrid}

\subsubsection*{\texttt{style=header-striped}}

\begin{verbatim}
\pxTABsetup
  {
    style=header-striped,
    headerbgcolor=blue!15,
    headertextcolor=blue!70!black,
    headerfont=\sffamily\bfseries,
    altrowbgcolor=gray!10,
    altrowtextcolor=black,
    tabwidth=.8\linewidth
  }
\pxTable{Name, Age, City}
        {Alice, 20, Barcelona}
        {Bob, 22, Madrid}
        {Clara, 23, Girona}
\end{verbatim}

\pxTABsetup
  {
    style=header-striped,
    headerbgcolor=blue!15,
    headertextcolor=blue!70!black,
    headerfont=\sffamily\bfseries,
    altrowbgcolor=gray!10,
    altrowtextcolor=black,
    tabwidth=.8\linewidth
  }
\pxTable{Name, Age, City}
        {Alice, 20, Barcelona}
        {Bob, 22, Madrid}
        {Clara, 23, Girona}

\subsection{Paragraph cells}

Paragraph mode produces equal-width columns that wrap long text,
suitable for rubrics, notes, and narrative content.  The following
example uses three columns with \texttt{cellmode=paragraph} and a
fixed table width:

\begin{verbatim}
\pxTABsetup
  {
    style=plain,
    cellmode=paragraph,
    paragraphcoltype=p,
    width=.8\linewidth
  }
\pxTable
  {Topic, Description, Status}
  {Design, The API was specified using key-value options
   to keep the interface consistent with existing pxTAB
   commands., Complete}
  {Tests, Paragraph regression tests exercise multi-line
   wrapping and interaction with all styles., Passed}
\end{verbatim}

\pxTABsetup
  {
    style=plain,
    cellmode=paragraph,
    paragraphcoltype=p,
    width=.8\linewidth
  }
\pxTable
  {Topic, Description, Status}
  {Design, The API was specified using key-value options
   to keep the interface consistent with existing pxTAB
   commands., Complete}
  {Tests, Paragraph regression tests exercise multi-line
   wrapping and interaction with all styles., Passed}

Paragraph mode works with all styles.  The next example combines
\texttt{header-striped} with vertically-centred \texttt{m} columns:

\begin{verbatim}
\pxTABsetup
  {
    style=header-striped,
    paragraphcells=true,
    paragraphcoltype=m,
    width=.85\linewidth,
    headerbgcolor=blue!12,
    headertextcolor=blue!70!black,
    altrowbgcolor=gray!10,
    altrowtextcolor=black
  }
\pxTable{Criterion, Comment}
  {Readability, This rubric comment wraps inside
   a paragraph cell while striped backgrounds fill
   the entire cell area across all wrapped lines.}
  {Correctness, Another multi-line comment to show
   that composed styles and paragraph columns work
   together seamlessly.}
\end{verbatim}

\pxTABsetup
  {
    style=header-striped,
    paragraphcells=true,
    paragraphcoltype=m,
    width=.85\linewidth,
    headerbgcolor=blue!12,
    headertextcolor=blue!70!black,
    altrowbgcolor=gray!10,
    altrowtextcolor=black
  }
\pxTable{Criterion, Comment}
  {Readability, This rubric comment wraps inside
   a paragraph cell while striped backgrounds fill
   the entire cell area across all wrapped lines.}
  {Correctness, Another multi-line comment to show
   that composed styles and paragraph columns work
   together seamlessly.}

\subsection{Centered and floating tables}

The new \texttt{centering} and float-related keys can be combined with
any of the predefined styles.  The following example uses
\texttt{style=header} with global centering:

\begin{verbatim}
\pxTABsetup
  {
    style=header,
    headerbgcolor=blue!15,
    headertextcolor=blue!70!black,
    headerfont=\sffamily\bfseries,
    centering=true
  }
\pxTable{Name, Age, City}
        {Alice, 20, Barcelona}
        {Bob, 22, Madrid}
\end{verbatim}

\pxTABsetup
  {
    style=header,
    headerbgcolor=blue!15,
    headertextcolor=blue!70!black,
    headerfont=\sffamily\bfseries,
    centering=true
  }
\pxTable{Name, Age, City}
        {Alice, 20, Barcelona}
        {Bob, 22, Madrid}

A float table with caption and label can be created by passing the new
keys in the options argument:

\begin{verbatim}
\pxTable[
  style=header,
  caption={Sample grades},
  label=tab:grades,
  float=htbp
]{Name, Grade}
 {Alice, 9}
 {Bob, 7}
\end{verbatim}

\pxTable[
  style=header,
  caption={Sample grades},
  label=tab:grades,
  float=htbp
]{Name, Grade}
 {Alice, 9}
 {Bob, 7}

As usual, the table can later be referenced with
\verb|Table~\ref{tab:grades}| in the surrounding text.

\section{Command reference}

This section summarises the user-facing commands provided by
\texttt{pxTAB}.  All commands respect the key settings described
above.

\subsection{\texttt{\textbackslash pxTABsetup}}

\begin{verbatim}
\pxTABsetup{<key>=<value>,...}
\end{verbatim}

Sets package-wide defaults in the \texttt{pxTAB} key family.
Usual \LaTeX\ scoping rules apply, so placing a setup call inside a
group affects only the tables in that group.

\subsection{\texttt{\textbackslash pxTable}}

\begin{verbatim}
\pxTable[<options>]{<row1>}{<row2>}...{<rowN>}
\end{verbatim}

Builds a table from braced row arguments.  Each \verb|<row>| is a
comma-separated list of cell contents.  A typical example is:

\begin{verbatim}
\pxTABsetup{style=header}
\pxTable{Name, Age, City}
        {Alice, 20, Barcelona}
        {Bob, 22, Madrid}
\end{verbatim}

\subsection{\texttt{\textbackslash pxTableFromList}}

\begin{verbatim}
\pxTableFromList[<options>]{<data>}
\end{verbatim}

Builds a table from a single list-style string.  By default rows are
separated by semicolons (\verb|;|) and cells by commas (\verb|,|):

\begin{verbatim}
\pxTableFromList
  {Name, Age, City;
   Alice, 20, Barcelona;
   Bob, 22, Madrid}
\end{verbatim}

The separators can be customised with the \texttt{rowsep} and
\texttt{cellsep} keys.

\subsection{\texttt{\textbackslash pxTableRow}}

\begin{verbatim}
\pxTableRow[<options>]{<row>}
\end{verbatim}

Convenience wrapper for single-row tables (headers or small data
rows).  It behaves as:

\begin{verbatim}
\pxTableRow[<opts>]{<row>} == \pxTable[<opts>]{<row>}
\end{verbatim}

\subsection{\texttt{\textbackslash pxTableHeadBody}}

\begin{verbatim}
\pxTableHeadBody[<options>]{<header>}{<body-list>}
\end{verbatim}

The header row is given explicitly as a comma-separated list, while
the body is provided as list-style data using the current separators:

\begin{verbatim}
\pxTableHeadBody
  {Name, Role}
  {Alice, Dev; Bob, QA; Carol, PO}
\end{verbatim}

\subsection{\texttt{\textbackslash pxTableHeadBodyRows}}

\begin{verbatim}
\pxTableHeadBodyRows[<options>]
  {<header>}{<row1>}{<row2>}...
\end{verbatim}

Variant of the previous command where both header and body rows are
given as separate braced arguments:

\begin{verbatim}
\pxTableHeadBodyRows
  {Name, Score}
  {Anna, 8}
  {Marc, 9}
  {Joan, 10}
\end{verbatim}

\subsection{\texttt{\textbackslash pxTableMatrix}}

\begin{verbatim}
\pxTableMatrix[<options>]{<rows>}{<cols>}{<flat-cells>}
\end{verbatim}

Builds a table from a flat list of cells, filled row by row.  Missing
cells are left blank:

\begin{verbatim}
\pxTableMatrix{2}{3}
  {m11,m12,m13,
   m21,m22,m23}
\end{verbatim}

\subsection{Quick reference}

The following table summarises the most common commands and their
arguments; all of them honour the global and per-table keys described in
the configuration section.

\begin{center}
\begin{tabular}{llp{.45\linewidth}}
\textbf{Command} & \textbf{Arguments} & \textbf{Typical use}\\\hline
\verb|\pxTABsetup| & \verb|{<key>=<value>,...}| &
  Set global defaults (style, width, centering, colours, float keys).\\
\verb|\pxTable| &
  \verb|[<options>]{<row1>}{...}{<rowN>}| &
  Small to medium tables entered as explicit rows.  Supports
  \texttt{centering}, \texttt{caption}, \texttt{label}, \texttt{float}.\\
\verb|\pxTableFromList| &
  \verb|[<options>]{<data>}| &
  Tables given as a single list string separated by \texttt{rowsep} and
  \texttt{cellsep}.\\
\verb|\pxTableRow| &
  \verb|[<options>]{<row>}| &
  Single-row tables (headers or quick inline tables).\\
\verb|\pxTableHeadBody| &
  \verb|[<options>]{<header>}{<body-list>}| &
  Head/body split with the body given as list-style data.\\
\verb|\pxTableHeadBodyRows| &
  \verb|[<options>]{<header>}{<row1>}{...}| &
  Head/body split where both header and body rows are given explicitly.\\
\verb|\pxTableMatrix| &
  \verb|[<options>]{<rows>}{<cols>}{<flat-cells>}| &
  Matrix-style data where a flat cell list is reshaped into a grid.\\
\end{tabular}
\end{center}

The float-related keys \texttt{caption}, \texttt{label}, and
\texttt{float}, together with \texttt{centering}, can be provided either
globally via \verb|\pxTABsetup| or per table in the \verb|[<options>]|
argument.

\section{Advanced usage}

\subsection{Best practices}

\begin{itemize}
  \item Use \texttt{style=plain} for compact, in-text tables where visual
    emphasis is not required; \texttt{style=header} works well for small
    datasets with a clear header row.
  \item For longer datasets, \texttt{style=striped} improves readability,
    while \texttt{boxed} and \texttt{grid} are suitable when a strong
    tabular structure is desired (summaries, rubrics, code lists).
  \item Prefer the natural-width default together with
    \texttt{centering=true} when the table is narrower than the text
    block; this keeps columns tight while visually centring the table.
  \item Use \verb|width=\linewidth| (or a fraction thereof) when the
    table should align with surrounding paragraphs; in this case
    \texttt{centering} is optional but harmless.
  \item When using floats, set both \texttt{caption} and \texttt{label}
    so that tables integrate with the document’s list of tables and can
    be referenced with \verb|\ref|.
\end{itemize}

\subsection{When to use paragraph mode}

Paragraph mode (\texttt{cellmode=paragraph}) is designed for tables
whose cells contain narrative text, rubric comments, or any content
that benefits from line wrapping:

\begin{itemize}
  \item \textbf{Prefer paragraph mode} when one or more columns hold
    multi-line descriptions, notes, or explanatory text.  Typical
    examples include rubric tables, requirement matrices, and
    coursework summary tables.
  \item \textbf{Prefer inline mode} (the default) when all columns
    contain short labels, numbers, or single-line entries.  Inline
    columns produce tighter, more compact tables.
  \item \textbf{Width interaction.}  Paragraph mode always requires an
    effective width to compute per-column dimensions.  When
    \texttt{width=natural} (the default), \verb|\linewidth| is used
    automatically; for narrower tables set an explicit width such as
    \verb|width=0.8\linewidth|.  The \texttt{centering} key works
    identically in paragraph and inline mode---it controls horizontal
    placement of the table block, not the column widths.
  \item \textbf{Vertical alignment.}  Use \texttt{paragraphcoltype=p}
    (default) for top-aligned cells or \texttt{paragraphcoltype=m}
    when rows of unequal height should be vertically centred.
\end{itemize}

\paragraph{Known limitations.}
\begin{itemize}
  \item Paragraph mode produces \emph{equal-width} columns only;
    heterogeneous widths require a manual \texttt{colspec}.
  \item An explicit \texttt{colspec} overrides the automatic paragraph
    layout entirely---users who set \texttt{colspec} are responsible
    for choosing appropriate \texttt{p\{...\}} or \texttt{m\{...\}}
    column types.
  \item There is no \texttt{longtable} integration; page-breaking
    paragraph tables remain the domain of \texttt{pxTBL}'s
    \verb|\pxTBLlong| command or manual \texttt{longtable} usage.
\end{itemize}

\subsection{Explicit column specifications}

When \texttt{colspec} is set, its value is used directly as the
\texttt{tabular} column specification:

\begin{verbatim}
\pxTABsetup{style=plain,colspec=ll}
\pxTable{Col 1, Col 2}
        {Left, also left}
\end{verbatim}

Simple \texttt{l}/\texttt{c}/\texttt{r} specifications are checked
against the inferred column count; on mismatch a warning is issued
and an automatic specification is used.

\subsection{Width control}

The \texttt{width}/\texttt{tabwidth} keys allow fixed-width tables:

\begin{verbatim}
\pxTABsetup{style=plain,width=\linewidth,align=l}
\pxTable{A, B, C}{1, 2, 3}{4, 5, 6}
\end{verbatim}

Non-\texttt{natural} widths use a \texttt{tabular*} layout with
automatic extra column separation so that the table fits the requested
width.

\subsection{Font size and row spacing}

\begin{verbatim}
\pxTable[fontsize=\small,rowstretch=1.05,style=header]
        {Name, Notes}
        {Alice, Compact rows with smaller text}
        {Bob, This checks per-table sizing controls}
\end{verbatim}

The \texttt{fontsize}/\texttt{size} and \texttt{arraystretch}/%
\texttt{rowstretch} keys apply only to the current table.

\subsection{Troubleshooting}

\begin{itemize}
  \item \emph{Empty or missing tables.}  If the first row has no cells,
    \texttt{pxTAB} skips the table and issues a warning.  Check that
    your header row is non-empty and that the separators
    (\texttt{rowsep}, \texttt{cellsep}) are correct.
  \item \emph{Column specification mismatch.}  When \texttt{colspec} is
    given but its length does not match the inferred column count, an
    automatic specification is used instead and a warning is written to
    the log.  Adjust \texttt{colspec} or let \texttt{pxTAB} derive it
    from \texttt{align}.
  \item \emph{Float placement warnings.}  The \texttt{float} key is
    passed directly to \verb|\begin{table}[...]|.  If you use an
    unsupported placement specifier, \LaTeX\ may report an ``Unknown
    float option'' warning; choose a standard value such as
    \texttt{h}, \texttt{t}, or \texttt{htbp}.
  \item \emph{Missing caption or label.}  A float without \texttt{label}
    still typesets correctly but cannot be referenced with
    \verb|\ref|.  Conversely, a \texttt{label} without \texttt{caption}
    is allowed; the table is still wrapped in a \texttt{table}
    environment so that the label has a proper float context.
\end{itemize}

\section{Implementation notes}

This subsection describes, at a high level, how \texttt{pxTAB} processes
table data and renders it.  It is intended for maintainers and advanced
users who need to understand the internal pipeline.

\subsection*{Row parsing and trailing empty trimming}

Input data (from \verb|\pxTable|, \verb|\pxTableFromList|, or the
head/body helpers) is stored as a sequence of rows in
\verb|\l__pxtab_rows_seq|.  For list-style input, rows are split using
the \texttt{rowsep} separator (default \texttt{;}), and cells within each
row using \texttt{cellsep} (default \texttt{,}).

Before rendering, \verb|\__pxtab_trim_trailing_empty_rows:| recursively
removes trailing rows that have no non-blank cells.  A row is considered
empty if every cell, after splitting by \texttt{cellsep}, is blank.
This ensures that trailing semicolons (e.g.\ \texttt{A,B;C,D;}) do not
produce extra blank rows in the output.

\subsection*{Column count and colspec derivation}

The column count is taken from the \emph{first} row only: it is the
number of cells after splitting by \texttt{cellsep}.  If the first row
is empty, a warning is emitted and the table is skipped.

The column specification (\texttt{colspec}) is determined as follows:
\begin{itemize}
  \item If the user provided \texttt{colspec} and it is a simple
    \texttt{l}/\texttt{c}/\texttt{r} string whose length matches the
    inferred column count, it is used as-is.
  \item If the provided \texttt{colspec} length does not match, a
    \texttt{colspec-mismatch} warning is issued and an automatic
    colspec is built from the \texttt{align} key.
  \item If no \texttt{colspec} was given, one is built automatically
    by repeating the \texttt{align} character (default \texttt{c}) for
    each column.
\end{itemize}

For \texttt{boxed} and \texttt{grid} styles, vertical rules are added
around the colspec (e.g.\ \texttt{|c|c|c|}).

\subsection*{Style renderers and shared helpers}

Each style has a dedicated renderer
(\verb|\__pxtab_render_table_plain:n|, \verb|header:n|,
\verb|striped:n|, \verb|boxed:n|, \verb|grid:n|).  All renderers share:

\begin{itemize}
  \item \verb|\__pxtab_append_row_to_body:nnnn| --- appends a row of
    cells to the body token list, with optional text colour and font.
  \item \verb|\__pxtab_append_row_styled_to_body:nnnnn| --- same, plus
    optional \verb|\rowcolor| for background.
  \item \verb|\__pxtab_emit_tabular_with_body:N| --- applies visual
    setup (\texttt{tabcolsep}, \texttt{arraystretch}, \texttt{rulecolor},
    etc.) and emits the \texttt{tabular} or \texttt{tabular*} with the
    accumulated body.
\end{itemize}

Style-specific behaviour:
\begin{itemize}
  \item \texttt{plain}: all rows use body font; no row colours or rules.
  \item \texttt{header}: first row gets \texttt{headerbgcolor},
    \texttt{headertextcolor}, \texttt{headerfont}; body rows use body font.
  \item \texttt{striped}: odd rows plain; even rows get
    \texttt{altrowbgcolor} and \texttt{altrowtextcolor}.
    Backgrounds are applied via \verb|\rowcolor| so they fill the
    entire cell area (including wrapped lines in paragraph mode),
    not just the text extent.
  \item \texttt{boxed}: colspec wrapped with \texttt{|}; \texttt{\textbackslash hline}
    before first row and after last row.
  \item \texttt{grid}: colspec wrapped with \texttt{|}; \texttt{\textbackslash hline}
    before and after every row.
  \item \texttt{header-boxed}: header-style first row (colours and font)
    with boxed frame rules as in \texttt{boxed}.
  \item \texttt{header-grid}: header-style first row together with full
    grid rules as in \texttt{grid}.  The header row background
    uses \verb|\rowcolor|, filling the entire cell area.
  \item \texttt{header-striped}: header-style first row; subsequent body
    rows alternate plain and striped colours like \texttt{striped}, with
    the striping parity starting at the first body row.  As with
    \texttt{striped}, row backgrounds fill the full cell width across
    all columns and wrapped lines.
\end{itemize}

\section{Integration with P3CTeX}

When using the \texttt{P3CTeX} document class and its core package
\texttt{pxCORE}, the \texttt{pxTAB} package is loaded as part of the
default configuration so that table commands are available without
extra setup.

\subsection{Branded presets}

When \texttt{pxTAB} is loaded via the \texttt{P3CTeX} class, five
branded presets are automatically registered before package option
processing, so they can even be used in \verb|\usepackage| options:

\begin{verbatim}
\usepackage[preset=p3c-header]{pxTAB}
\end{verbatim}

The branded presets use the P3CTeX colour palette defined in
\texttt{pxGDX} (\texttt{primaryColor}, \texttt{darkColor},
\texttt{secondaryColor}, etc.).

\begin{center}
\begin{tabular}{lp{.6\linewidth}}
\textbf{Preset} & \textbf{Description}\\\hline
\texttt{p3c-header} &
  Standard branded header table: deep navy header with white text
  (\texttt{style=header}, sans-serif bold, centred).\\
\texttt{p3c-data} &
  Striped data table with branded colours: navy header plus light
  cyan alternating rows (\texttt{style=header-striped}).\\
\texttt{p3c-rubric} &
  Assessment rubric: full grid with paragraph cells, dark blue header
  (\texttt{style=header-grid}, \texttt{cellmode=paragraph}).\\
\texttt{p3c-minimal} &
  Clean minimal table: no header styling, tight spacing, neutral
  rule colour (\texttt{style=plain}).\\
\texttt{p3c-exam} &
  Exam/assessment table: dark blue header with light grey alternating
  rows, serif bold header, slightly wider row spacing
  (\texttt{style=header-striped}).\\
\end{tabular}
\end{center}

\noindent Example:

\begin{verbatim}
\documentclass{P3CTeX}
% pxTAB is loaded automatically; branded presets are available.

\begin{document}
\pxTable[preset=p3c-header]{Name, Grade}{Alice, 9}{Bob, 7}
\end{document}
\end{verbatim}

\subsection{Standalone loading}

When \texttt{pxTAB} is loaded outside the \texttt{P3CTeX} class (e.g.\
with \verb|\documentclass{article}|), the branded presets are
\emph{not} registered.  The package behaves identically to a standalone
installation: all generic built-in presets (\texttt{formal},
\texttt{zebra}, \texttt{minimal}) remain available, and users can
define their own presets with \verb|\pxTABsavepreset|.

\section{Legacy pxTBL commands}

The three commands described below provide a positional-argument
interface to paragraph-column tables.  They predate the key--value API
of \texttt{pxTAB} and are retained for backward compatibility.  For
full key--value control, use the \texttt{pxTAB} commands instead.

All three commands accept a star (\texttt{*}) variant that adds
alternating row colours (hard-coded \texttt{blue!25!}).

\subsection{\texttt{\textbackslash pxTBLtop}}

\begin{verbatim}
\pxTBLtop*[<maxwidth>]{<numcols>}[<gap>]{<headers>}[<prefix>]{<body>}
\end{verbatim}

Builds a fixed-position table (inside a \texttt{table} float) with
equal-width paragraph columns.  The \verb|<headers>| and \verb|<body>|
arguments contain \emph{raw tabular content} (cells separated by
\texttt{\&}, rows by \verb|\\|).

\begin{itemize}
  \item \texttt{*} --- starred variant adds \verb|\rowcolors| for
    alternating rows.
  \item \texttt{[<maxwidth>]} --- total table width (default
    \verb|\linewidth|).
  \item \texttt{\{<numcols>\}} --- number of columns.
  \item \texttt{[<gap>]} --- inter-column separator passed as
    \verb|@{<gap>}| (default \verb|\quad|).
  \item \texttt{\{<headers>\}} --- raw tabular header row.
  \item \texttt{[<prefix>]} --- column type: \texttt{p} (top-aligned)
    or \texttt{m} (vertically centred); default \texttt{p}.
  \item \texttt{\{<body>\}} --- raw tabular body rows.
\end{itemize}

\noindent Example:

\begin{verbatim}
\pxTBLtop{3}{Header A & Header B & Header C}
  {Row 1 Col A & Row 1 Col B & Row 1 Col C \\
   Row 2 Col A & Row 2 Col B & Row 2 Col C}
\end{verbatim}

\subsection{\texttt{\textbackslash pxTBLlong}}

\begin{verbatim}
\pxTBLlong*[<maxwidth>]{<numcols>}[<gap>]{<headers>}[<prefix>]{<body>}
\end{verbatim}

Builds a page-breaking \texttt{longtable} with the header row repeated
on every page.  Arguments are identical to \verb|\pxTBLtop|, but the
output uses \texttt{longtable} instead of a \texttt{table} float.

\noindent Example:

\begin{verbatim}
\pxTBLlong{2}{Name & Score}
  {Alice & 9 \\ Bob & 7 \\ Carol & 8}
\end{verbatim}

\paragraph{Known quirk.}  The \texttt{[<gap>]} optional argument is
accepted but \emph{silently ignored} by \verb|\pxTBLlong|---the
longtable colspec does not include \verb|@{<gap>}| separators.

\subsection{\texttt{\textbackslash pxTBL}}

\begin{verbatim}
\pxTBL*[<maxwidth>]{<numcols>}[<gap>]{<headers>}[<prefix>]{<body>}
\end{verbatim}

Convenience command that takes \emph{comma-separated lists} for
\verb|<headers>| and \verb|<body>| (instead of raw tabular content)
and delegates to the \texttt{longtable} renderer internally.  Each
body item becomes one row; items are separated by \verb|\hline|.

\noindent Example:

\begin{verbatim}
\pxTBL{3}{Name, Age, City}{Alice, 20, Barcelona, Bob, 22, Madrid}
\end{verbatim}

\paragraph{Known quirk.}  The \texttt{[<gap>]} argument is silently
ignored (inherited from the underlying \verb|\pxTBLlong| renderer).

\subsection{Star variant differences}

The star variant semantics differ slightly between commands:
\begin{itemize}
  \item \verb|\pxTBLtop*| colours \emph{odd} body rows only
    (\verb|\rowcolors*{1}{blue!25!}{}|).
  \item \verb|\pxTBLlong*| and \verb|\pxTBL*| colour \emph{all} rows
    (\verb|\rowcolors*{1}{blue!25!}{blue!25!}|).
\end{itemize}

\end{document}

